The state learner \\ {-- a super learner for right-censored data}

### 2024-05-10 tex/jrss-version/state-learner-jrss-style.tex
add event and censoring probabiltiy

In this section we report results from a simulation study. In the
first part we consider estimation of the conditional survival
function. Here we compare the state learner to two IPCW based discrete
super learners that use either the Kaplan-Meier estimator or a Cox
model to estimate the censoring probability
\cite{gonzalez2021stacked}.  In the second part we compare the state
learner to the super learner proposed by \cite{westling2021inference}.

@@

We compare the state learner to two IPCW based super learners: The
first superlearner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
% ANDERS: describe that you use a single evaluation time point and say how many censored, event happen before it.
Each super learner provides a learner for the cumulative hazard
function for the outcome of interest. From the cumulative hazard
function a risk prediction model can be obtained (c.f.,
equation~(\ref{eq:surv-def}) with $\Lambda_2 = 0$). We measure the
performance of each super learner by calculating the index of
prediction accuracy (IPA) \citep{kattan2018index} at a fixed time
horizon (36 months) for the risk prediction model provided by the
super learner. The IPA is 1 minus the ratio between the model's Brier
score and the null model's Brier score, where the null model is the
model that does not use any covariate information. The IPA is
approximated using a large (\( n = 20,000 \)) independent data set of
uncensored data. As a benchmark we calculate the performance of the
risk prediction model chosen by the oracle selector, which uses the
large data set of uncensored event times to select the learner with
the highest IPA.

>>>>

In this section we report results from a simulation study. In the first part we
consider estimation of the conditional survival function. Here we compare the
state learner to two IPCW based discrete super learners that use either the
Kaplan-Meier estimator or a Cox model to estimate the censoring probability
\citep{gonzalez2021stacked}. In the second part we compare the state learner to
the super learner proposed by \cite{westling2021inference}.

@@

We compare the state learner to two IPCW based super learners: The
first superlearner, called IPCW(Cox), uses a Cox model with main
effects to estimate the censoring probabilities, while the second
super learner, called IPCW(KM), uses the Kaplan-Meier estimator to
estimate the censoring probabilities. The Cox model for the censoring
distribution is correctly specified in both simulation settings while
the Kaplan Meier estimator only estimates the censoring model
correctly in the simulation setting where censoring is
independent. Both IPCW super learners are fitted using the
\texttt{R}-package \texttt{riskRegression}
\citep{Gerds_Ohlendorff_Ozenne_2023}.
% ANDERS: describe that you use a single evaluation time point and say how many
% censored, event happen before it.
%
% THOMAS: Done below now, OK?
%
% | time | sim_setting | true_events | true_cens | at_risk |
% |------+-------------+-------------+-----------+---------|
% |   36 | original    |      24.619 |    61.853 |  25.774 |
% |   36 | indep_cens  |      24.674 |    38.740 |  46.141 |
%
The IPCW super learners use the integrated Brier score up to a fixed time
horizon (36 months). The marginal risk of the event before this time horizon is
25\%. Under the ``dependent censoring'', setting the marginal censoring
probability before this time horizon is 62\%; under the ``independent
censoring'' setting, the marginal censoring probability before this time horizon
is 39\%.

Each super learner provides a learner for the cumulative hazard function for the
outcome of interest. From the cumulative hazard function a risk prediction model
can be obtained (c.f., equation~(\ref{eq:surv-def}) with $\Lambda_2 = 0$). We
measure the performance of each super learner by calculating the index of
prediction accuracy (IPA) \citep{kattan2018index} at a fixed time horizon (36
months) for the risk prediction model provided by the super learner. The IPA is
1 minus the ratio between the model's Brier score and the null model's Brier
score, where the null model is the model that does not use any covariate
information. The IPA is approximated using a large (\( n = 20,000 \))
independent data set of uncensored data. As a benchmark we calculate the
performance of the risk prediction model chosen by the oracle selector, which
uses the large data set of uncensored event times to select the learner with the
highest IPA.
